PMC Survey Data Analysis

### 2021-08-08 analysis.Rmd
scoring scheme

@@

@@

get_cites <- function(PMCID) {
  mylink <- elink(uid=PMCID , dbFrom="pmc", dbTo = "pubmed")
  PMID <- mylink[1]
  CITES <- elink(uid = PMID , dbFrom = "pubmed"  , linkname = "pubmed_pubmed_citedin" )
  N_CITES <- length(CITES[1])
  PUBDATE <- get_pub_date(PMCID)
  return(c(PMCID,PMID,N_CITES,PUBDATE))
}

PMCIDS <- as.numeric(gsub("PMC","",x$Pubmed.Central.ID))

# getcites needs to be fixed so that it can handle all the results
PMCIDS <- head(PMCIDS,100)
res <- lapply(X = PMCIDS, FUN = get_cites)
mx <- do.call(rbind,res)
mx <- apply(X = mx, MARGIN = 2, FUN = as.numeric)
df <- as.data.frame(mx)
colnames(df) <- c("PMCID","PMID","NumCites","Pubdate")

@@

mylm

>>>>

## Score

Here I am proposing a scoring scheme.
If there is information missing from the article, or basic mistakes are made, then a point is deducted.
If the article goes over and above the basic reproducibility, then they are awarded a point.

Gene set library origin not stated = -1

Gene set library version not stated = -1

Stat test not stated = -1

No stat test conducted = -1

No FDR correction conducted = -1

App used not stated = -1

App version not stated = -1

Background list not defined = -1

Inappropriate background list used = -1

Code availability = +1

Gene lists provided = +1

```{r,score}

head(x)
x1 <- x[1,]
x1

score <- function(r){
  r[is.na(r)] <- 0
  SCORE=0
  # gene set lib
  if(r[7]=="Not stated"){
    SCORE=SCORE-1
  }
  # GS version
  if(r[8]=="No"){
    SCORE=SCORE-1
  }
  # stat test
  if(r[9]=="No test"){
    SCORE=SCORE-1
  }
  if(r[9]=="Not stated"){
    SCORE=SCORE-1
  }
  # FDR
  if(r[10]!="Yes"){
    SCORE=SCORE-1
  }
  # app used
  if(r[11]=="Not stated"){
    SCORE=SCORE-1
  }
  # App version
  if(r[12]=="No"){
    SCORE=SCORE-1
  }
  # Code availability
  if(r[13]=="Yes"){
    SCORE=SCORE+1
  }
  # Background
  if(r[14]=="Not stated"){
    SCORE=SCORE-1
  }
  if(r[14]=="No"){
    SCORE=SCORE-1
  }
  if(r[14]=="Stated, but incorrect"){
    SCORE=SCORE-1
  }
  # gene list
  # Code availability
  if(r[16]=="Yes"){
    SCORE=SCORE+1
  }
  return(SCORE)
}

scores <- apply(X = x,MARGIN = 1, FUN = score)
hist(scores)
x$scores <- scores

@@

## Analysis scores vs Journal rank

```{r,scoresvsjrank}

colnames(xm)

tail(xm$Journal)

mylm1 <- lm (xm$scores ~ xm$SJR)
plot(xm$SJR,xm$scores,xlab="SJR",ylab="score",main="Association of analysis quality with journal metrics")
abline(mylm1,col="red")

cor.test(xm$scores,xm$SJR,method="pearson")

cor.test(xm$scores,xm$SJR,method="spearman")
```
## Best and worst journals

Minimum of 5 analyses.

```{r,goodbadugly}

colnames(xm)

xmm <-xm[,c(1,17)]
tail(xmm)
tab <- table(xmm$Journal)
tab <- tab[which(tab>5)]
xmm <- xmm[which(xmm$Journal %in% names(tab)),]
head(xmm)
dim(xmm)
jres <- aggregate(. ~ Journal, xmm ,mean)
jres <- jres[order(jres$scores),]
head(jres)
dim(jres)

par(mar=c(5,12,3,1))
barplot(tail(jres$scores,20),names.arg = tail(jres$Journal,20),horiz=TRUE,las=1,cex.names = 0.7, xlab="scores",
        main = "Best journals",xlim=c(-4,0))
grid()

par(mar=c(5,12,3,1))
barplot(head(jres$scores,20),names.arg = head(jres$Journal,20),horiz=TRUE,las=1,cex.names = 0.7, xlab="scores",
        main = "Best journals",xlim=c(-5,0))
grid()

## Mean analysis scores vs Journal rank

```{r,scoresvsjrank}

head(jres)
head(mjdf)

jjres <- merge(jres,mjdf,by.x="Journal",by.y="journalabbrev",all.x = TRUE)

head(jjres)

mylm2 <- lm (jjres$scores ~ jjres$SJR)
plot(jjres$SJR,jjres$scores,xlab="SJR",ylab="score",main="Association of mean analysis quality with journal metrics")
abline(mylm2,col="red")

cor.test(jjres$scores,jjres$SJR,method="pearson")

cor.test(jjres$scores,jjres$SJR,method="spearman")

@@

get_cites <- function(PMCID) {
  mylink <- elink(uid=PMCID , dbFrom="pmc", dbTo = "pubmed")
  PMID <- mylink[1]
  CITES <- elink(uid = PMID , dbFrom = "pubmed"  , linkname = "pubmed_pubmed_citedin" )
  N_CITES <- length(CITES[1])
  return(c(PMCID,PMID,N_CITES))
}

```{r,runcites}

if (file.exists("cites_data.rds") ) {
  res <- readRDS("cites_data.rds")
} else {
  PMCIDS <- as.numeric(gsub("PMC","",x$Pubmed.Central.ID))
  res <- lapply(X = PMCIDS, FUN = get_cites)
  saveRDS(res,file="cites_data.rds")
}
res
mx <- do.call(rbind,res)
mx <- apply(X = mx, MARGIN = 2, FUN = as.numeric)
df <- as.data.frame(mx)
colnames(df) <- c("PMCID","PMID","NumCites","Pubdate")

@@

mylm
df$Pubmed.Central.ID <- gsub("^","PMC",df$PMCID)
head(df)

dfm <- merge(df,xm,by.x=,by.y="Pubmed.Central.ID")
colnames(dfm)

mylm2 <- lm (dfm$scores ~ dfm$NumCites)
plot(dfm$NumCites,dfm$scores,xlab="Num cites",ylab="score",main="Association of citations with journal metrics")
abline(mylm2,col="red")
